---
title: "About the data"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
runtime: shiny
resource_files:
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.cpg
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.dbf
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.prj
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbn
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.sbx
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shp.xml
- Data/Shapefiles/County shapefiles/MNCounties_MNDOT.shx
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(sf)
library(ggrepel)
library(scales)
library(shiny)
library(shinycssloaders)
library(ggiraph)
library(kableExtra)
library(rmapshaper)
library(cowplot)
library(DT)
```

```{r loading jon docs and shapefiles, cache=TRUE, include=FALSE}
theme_bar <- theme_bw() +
  theme(panel.grid.major = element_line(color = "grey70", size = 0.1),
        panel.grid.minor = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        legend.title = element_blank(),
        text = element_text(size = 15),
        plot.caption = element_text(hjust = 0.5, face = "italic"),
        legend.text = element_text(margin = margin(l = 2, r = 5)))

theme_line <- theme_bw() +
  theme(legend.background = element_rect(fill = "transparent", color = "transparent"),
        legend.key = element_rect(fill = "transparent"),
        panel.grid.minor = element_blank(),
        panel.grid.major = element_line(color = "grey70", size = 0.1),
        axis.ticks = element_blank(),
        axis.text = element_text(face = "bold"),
        panel.border = element_blank(),
        text = element_text(size = 15),
        plot.caption = element_text(hjust = 0.5, face = "italic"),
        legend.text = element_text(margin = margin(l = 2, r = 5)))

theme_sf <- theme_bw() +
  theme(axis.text.x=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks=element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_line(color = "white"),
        panel.border = element_blank(),
        plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        text = element_text(size = 15),
        plot.caption = element_text(hjust = 0.5, face = "italic"),
        legend.text = element_text(margin = margin(l = 2, r = 5)))

counties.regions <- read_csv("Data/Join docs/county_regions.csv") %>%
  mutate(countyfp = formatC(countyfp, width = 3, flag = "0"),
         Name = str_to_title(Name),
         Name = str_replace(Name, "Q", "q"),
         Name = str_replace(Name, "Of The", "of the"),
         Dem_Desc = ifelse(Name == "Minnesota", "Minnesota", Dem_Desc),
         Name = str_replace(Name, "Mcleod", "McLeod"),
         planning.region = str_replace(planning.region, " Minnesota", ""),
         edr = str_replace(edr, "  ", " "))

edr.pr <- counties.regions %>%
  distinct(edr, .keep_all = TRUE) %>%
  select(5,6) %>%
  mutate(edr = str_replace(edr, "  ", " "),
         planning.region = str_replace(planning.region, " Minnesota", ""))

color.ruca <- c("Entirely rural" = "#5CA81F", "Town/rural mix" = "#C7EF99", "Urban/town/rural mix" = "#d8b365", "Entirely urban" = "#a6611a", "Minnesota" = "black")

color.pr <- c("Northwest" = "#810f7c","Northeast" = "#fe9929", "Central" = "#076324", "Seven County Mpls-St Paul" = "#d8b365", "Southwest" = "#1f78b4", "Southeast" = "#d7301f", "Minnesota" = "black")

color.edr <- c("EDR 1 - Northwest" = "#b3cde3", "EDR 2 - Headwaters" = "#8c96c6", "EDR 3 - Arrowhead" = "#fe9929", "EDR 4 - West Central" = "#8856a7", "EDR 5 - North Central" = "#810f7c", "EDR 6E- Southwest Central" = "#e5f5f9", "EDR 6W- Upper Minnesota Valley" = "#bdc9e1", "EDR 7E- East Central" = "#99d8c9", "EDR 7W- Central" = "#2ca25f", "EDR 8 - Southwest" = "#74a9cf", "EDR 9 - South Central" = "#0570b0", "EDR 10 - Southeast" = "#d7301f", "EDR 11 - 7 County Twin Cities" = "#d8b365", "Minnesota" = "black")

color.counties <- scale_color_brewer(palette = "Dark2",
                       guide = guide_legend(ncol = 3))

mn_counties <- st_read("Data/Shapefiles/county shapefiles/MNCounties_MNDOT.shp", quiet = TRUE) %>%
  ms_simplify(keep = .01, keep_shapes = TRUE)
```

<br>

The data used in the following visualizations are taken from the U.S. Census Bureau - American Community Survey (2017 5-year). It is important to understand the the values provided are an average of a 5-year period (2013-2017).

# Questions asked in the survey

These are the following questions asked which provided the data used;

*Question 8: At this house, apartment, or mobile home - do you or any member of this household own or use any of the following types of computer?*

- [ ] Desktop or laptop
- [ ] Smartphone
- [ ] Tablet or other portable wireless computer
- [ ] Some other type of computer

*Question 9: At this house, apartment or mobile home - do you or any member of this household have access to the internet?*

- [ ] Yes, by paying a cell phone company or internet service provider
- [ ] Yes, without paying a cell phone company or internet service provider
- [ ] No access to the internet at this house, apartment, or mobile home.

*Question 10: Do you or any member of this household have access to the internet using a -*

- [ ] Cellular data plan for a smart phone or other mobile device (yes, no)
- [ ] broadband (high speed) internet service such as cable, fiber optic, or DSL service installed in this household
- [ ] satellite internet service installed in this household
- [ ] dial-up internet service installed in this household.
- [ ] some other service.